---
title: "Code"
author: "LR"
date: '2022-06-29'
output: html_document
---
# PFAS exposure of humans, animals and the environment: a systematic meta-review map and bibliometric analysis  

# supplementary Material - data processing and data visualization

# Setup and load packages

```{r setup, include=TRUE}
library(tidyverse)
library(here)
library(stringr)
library(knitr)
library(formatR)
library(forcats)
library(ggplot2)
library(hrbrthemes) 
library(patchwork)
library(bibliometrix)
library(igraph)


knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

# Load data

Manually extracted data using Google Forms is stored in the "data" folder. This data is splitted in four difference .csv files:
1. **Main.csv** = main data sheet for the systematic map
2. **AMSTAR2.csv** = AMSTAR2 quality assessment data sheet
3. **PFAS_types.csv** = data sheet about PFAS types investigated in the included systematic reviews
4. **Species.csv** = data sheet on species investigated in the included systematic reviews

Bibliographic data records are exported from Scopus (including cited references field) in .bib format and locally saved as **scopus.bib** in the "data" folder.

```{r load data}
mdata <- read_csv(here("data", "Main.csv"))
spdata <- read.csv(here("data", "Species.csv"))
qdata <- read.csv(here("data", "AMSTAR2.csv"))
ptdata <- read.csv(here("data", "PFAS_types.csv"))
# change data format
ptdata <- ptdata %>% 
  select(Study_ID:PFAS_type) %>% 
  separate_rows(PFAS_type, sep = ', ')
  
```

